% Co-expression Networks
% [Keith Hughitt](khughitt@umd.edu)
% 2013/04/011

Overview
========
Going through some of the co-expression network papers and putting together
notes... Will have to think a little more about how much time I want to put
in to producing these notes.

Scale-free networks
===================
Many biological networks (including co-expression networks) are thought to
follow a [power law distribution](https://en.wikipedia.org/wiki/Power_law)^1.

For co-expression networks with genes as nodes, the degree distribution $p(k)$
for genes follows:

$$
p(k) \sim k^{-\gamma}
$$

where $k$ is the number of connections to other genes.

The exponent $\gamma$ determines how quickly the distribution decays,
for example:

### $\gamma = 3$
```{r powerlaw_plot1}
k = seq(1, 10, by=0.1)
plot(k, k^-2, type='l')
```

### $\gamma = 10$
```{r powerlaw_plot2}
plot(k, k^-10, type='l')
```


References
==========
1. Bin Zhang, Steve Horvath,   (2005) A General Framework For Weighted Gene co-Expression Network Analysis.  *Statistical Applications in Genetics And Molecular Biology*  **4**  [10.2202/1544-6115.1128](http://dx.doi.org/10.2202/1544-6115.1128)>
